---
title: "Global health research"
resource_files:
- s01.2-What is global health research.Rmd
- css/custom-learnr.css
- images/logo_dark.png
- images/logo.png
- images/scholartoppubs.png
runtime: shiny_prerendered
output:
  learnr::tutorial:
    css: css/custom-learnr.css
subtitle: Session 1.2
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

![](https://youtu.be/2gPcbRvFyoQ)

Research is defined as a systematic investigation whose goal is to develop or contribute to generalizable knowledge. As global health is the study and practice of improving health and achieving equity in health, global health research research includes all systematic efforts to create generalizable knowledge of how we can improve health and achieve equity in health for everyone, everywhere.

Given the complexity of our global health challenges, global health research is multidisciplinary and interdisciplinary. This means that, as a budding global health researcher, you will likely collaborate on teams with colleagues from many different fields. Each discipline has its own methodological preferences and core approaches, but we all operate in the same research landscape that is divided into basic and applied research. Much of the literature in global health focuses on our applied challenges, so you will read a lot about clinical research and trials, translational research, implementation research, and M&E (monitoring and evaluation).

So no matter what discipline you call home, or what stage of research you find most interesting, you can make great contributions to our collective efforts to improve health and achieve equity in health.

### Learning Objectives

* Understand the definition of global health and what distinguishes global health from other related fields
* Explain what makes research scientific
* Describe the landscape of global health research
* Enumerate the phases of clinical trials
* Describe the key actors in the funding and production of global health research/initiatives
* List several scholarly journals that publish global health research

## Before class

### Prepare

* Read [GHR Chapter 1](http://read.themethodsection.com/ghr.html){target="_blank"}.
* Read ["Fighting glioblastoma with recombinant poliovirus"](https://medium.com/@ericpgreen/fighting-glioblastoma-with-recombinant-poliovirus-5bef1ebc7f41)
* Think about where your interests lie in the global health research landscape and be prepared to share with classmates.
* Browse one or more of the journals that publish global health research and find at least one article that interests you.

### Pre-Class Assessment

```{r pca01-2, echo=FALSE}
quiz(
question("Which of the following is NOT a main characteristic of scientific research that apply to most of global health?",
  answer("The conclusions are uncertain", message = "There are three main characteristics of scientific research that apply to most of global health, ‘the conclusions are uncertain’ is one of them."),
  answer("The goal is inference", message = "There are three main characteristics of scientific research that apply to most of global health, ‘the goal is inference’ is one of them."),
  answer("The procedures are private", correct = TRUE),
  allow_retry = TRUE
  ),
question("The goal to infer that some program or some treatment, intervention, or policy had an effect on some outcome is called _______.",
  answer("Impact evaluation", correct = TRUE),
  answer("Theory failure", message = "Theory of failure refers to the failure of a program because the idea or theory behind the program is wrong."),
  answer("Implementation failure", message = "Implementation failure refers to the troubled implementation of a program so much so that there was never a chance for the program to succeed."),
  allow_retry = TRUE
  ),
question("Determining a safe dosing range and evaluating the possible side effects of a drug are conducted in which phase of a clinical trial?",
  answer("Phase I", correct = TRUE),
  answer("Phase II", message = "Phase II of clinical trials studies the effects optimal dosing of the drug has on clinical markers and health outcomes."),
  answer("Phase III", message = "Phase III of clinical trials studies the effect of the drug on clinical markers or health outcomes when compared to an existing treatment or placebo in a randomized evaluation."),
  allow_retry = TRUE
  ),
question("Generation of hypotheses and theories from specific observations is called _______.",
  answer("Empiricism", message = "Empiricism is a more general term that is defined as using direct and indirect observations to draw conclusions."),
  answer("Deductive reasoning", message = "Deductive reasoning is generation of specific conclusions from general theories"),
  answer("Inductive reasoning", correct = TRUE),
answer("None of the above"),
  allow_retry = TRUE
  ),
question("Which of the following is an example of applied research?",
  answer("Observing the cellular mechanisms of skin repair in humans and other mammals", message = "This is a basic research to understand the mechanisms."),
  answer("Genetic modification of mosquitoes as a potential tool for elimination of malaria", correct = TRUE),
  answer("Studying the influences of dietary habits on taste perception and oral somatic sensation", message = "This is an example of basic research to understand the effects of dietary habits"),
answer("Studying the relationships between environmental factors and violent behaviors", message = "This is an example of basic research in psychology to understand what drives violent behaviors"),
  allow_retry = TRUE
  )
)
```


<!--
<iframe height="800" width="100%" src="https://duke.qualtrics.com/jfe/form/SV_bBg1PTRLi5Usjat" frameborder="0"></iframe>
-->

## During class

### Mini Lecture

![](https://youtu.be/kKQSroMLkWI)

[Deck](https://docs.google.com/presentation/d/e/2PACX-1vR94D7NV3OUFYfzZH8YX5ggN6LbuZ27z8lrW5LNYNtnEXIikcyazNuK4Vgy2ATATBH2QglY47_I1Aan/pub?start=false&loop=false&delayms=15000){target="_blank"} / [Downloadable full resolution screencast](https://www.dropbox.com/s/uj99dhcdf68eyed/GHR%20CH01.mp4?dl=0)

### Activity

In this activity you will consider where your interests lie in the global health research landscape and identify academic journals that publish work related to these interests.

<iframe src="https://drive.google.com/file/d/1kPIiAOMpilS-NfAJnqTc3z4kkecirnzg/preview" width="100%" height="480"></iframe>

[Download activity sheets](https://drive.google.com/open?id=1VKNGv6buP9iGLPEbrbdYsw1YtcKGNkdE)

## After class

An optional activity is to get to know your local global health resources. For students living in the Research Triangle, this includes:

* [Duke Global Health Institute](https://globalhealth.duke.edu/){target="_blank"}
* [UNC Gillings School of Global Public Health](https://sph.unc.edu/){target="_blank"}
* [fhi360](https://www.fhi360.org/){target="_blank"}
* [IntraHealth International](https://www.intrahealth.org/){target="_blank"}
* [RTI International](https://www.rti.org/){target="_blank"}

Visit the websites, take note of upcoming events, sign up for newsletters, and browse faculty/researcher profiles.
