---
title: "Developing research ideas"
resource_files:
- s01.3-Developing research ideas.Rmd
- css/custom-learnr.css
- images/logo_dark.png
- images/logo.png
runtime: shiny_prerendered
output:
  learnr::tutorial:
    css: css/custom-learnr.css
subtitle: Session 1.3
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

![](https://youtu.be/osrlbjbzntc)

Research ideas take time to develop. You can jumpstart the process by learning about interesting work. Attend talks, watch webinars, meet with mentors, talk with fellow students. Above all else, read the literature. When you are just starting out, try skimming widely read journals to get a sense of current developments in global health disciplines. Follow interesting leads by noting article keywords, inspecting reference lists, and seeking out clues authors leave you in the introduction and discussion.

Once you narrow your focus a bit, try searching scientific databases. If you wade into the open web, watch out for predatory journals masquerading as legitimate science (and hit delete on emails inviting you to publish your work, esteemed ladies and gentlemen).

And please, please use a reference manager to keep track of all the cool science you are reading.

### Learning Objectives

* Appreciate the process for developing research ideas and identifying potential research problems
* Conduct an effective literature search
* Understand options for accessing journal articles
* Use a reference manager

## Before class

### Prepare

* Read [GHR Chapter 2](https://read.themethodsection.com/ideas.html){target="_blank"}.
* Download and install [Zotero](https://www.zotero.org/){target="_blank"}

<!--
<iframe height="800" width="100%" src="https://duke.qualtrics.com/jfe/form/SV_bBg1PTRLi5Usjat" frameborder="0"></iframe>
-->

### Pre-Class Assessment

```{r pca01-3, echo=FALSE}
quiz(
question("Have you downloaded and installed Zotero on your computer?",
  answer("No", message = "Follow this link to download and install Zotero on your computer, https://www.zotero.org "),
  answer("Yes", correct = TRUE),
  allow_retry = TRUE
  ),
question("A good example of a research problem must be able to do which of the following?",
  answer("use non-public methods to gather and analyze data on the problem ", message = "A good research problem must use systematic, public methods to gather and analyze data to solve the problem of interest."),
  answer("use systematic, public methods to gather and analyze data on the problem", correct = TRUE),
  answer("use random, private methods to gather and analyze data on the problem ", message = "A good research problem must use systematic, public methods to gather and analyze data to solve the problem of interest."),
  allow_retry = TRUE
  ),
question("The “C” in the mnemonic PICO stands for?",
  answer("Competition", message = "PICO stands for P:Patient, Population, or   Problem, I: Intervention, Prognostic Factor, or Exposure, C:Comparison, O: Outcome."),
  answer("Confounder", message = "PICO stands for P:Patient, Population, or Problem, I: Intervention, Prognostic Factor, or Exposure, C:Comparison, O: Outcome."),
  answer("Comparison", correct = TRUE),
  allow_retry = TRUE
  ),
question("Which of the following is an example of good PICO research question?",
  answer("Among pregnant women living in malaria-endemic areas, is chemoprevention more effective than a placebo at preventing parasitemia?", correct = TRUE),
  answer("Among pregnant women living in malaria-endemic areas, is chemoprevention more effective than a placebo?", message = "This research question is missing the “outcome” term. What is the target? The research question does not answer what the chemoprevention is more effective at compared to the placebo."),
  answer("Is chemoprevention more effective than a placebo at preventing parasitemia?", message = "This research question is missing the population of interest.The question is not well focused to answer, “in what population is the study being conducted?”"),
  allow_retry = TRUE
  ),
question("A regular google search is a great way for searching the scientific literature",
  answer("True", message = "Studies on the open web may come from  open-access, academic publishing business models that involve charging publication fees to authors without checking articles for quality and legitimacy"),
  answer("False", correct = TRUE),
  allow_retry = TRUE
  )
)
```


## During class

### Mini Lecture

![](https://youtu.be/_ypKYJO_tGY)

[Deck](https://docs.google.com/presentation/d/e/2PACX-1vTRpncIaLcJQ2c--be9Id6OV72iOdTP1qt4R6OK7WfUJBJK5uUJpUBl7XXSEFtp0iC4AYhOt4aegAyZ/pub?start=false&loop=false&delayms=15000) / [Downloadable full resolution screencast](https://www.dropbox.com/s/yc3f81icdwn0jfd/GHR%20CH02.mov?dl=0)

### Activity

During class we will participate in a workshop led by Duke librarians that will teach you how to construct a research question, design and implement a search strategy, import the results into Zotero, and automatically compile and format bibliographies in a word processor.

![](https://youtu.be/IUh-ZG4JaLI)

[Deck](https://drive.google.com/open?id=1xlSYpLBRcqsHTmYViIJ17eKi6KePX-TA) / [Downloadable full resolution video](https://www.dropbox.com/s/tih6cwzxo0kcsfo/GHR%20Research%20Database%20and%20Zotero%20Workshop.mp4?dl=0)

## After class

An optional activity is to take your Zotero knowledge to the next level with [these learning resources](https://research.library.gsu.edu/c.php?g=115275&p=750658) from Georgia State University. Consider watching this short video on creating shared Zotero groups.

![](https://www.youtube.com/watch?v=Bag6XdJIPLw)
