---
title: "The role of theory in global health"
resource_files:
- s02.2.Rmd
- css/custom-learnr.css
- images/logo_dark.png
- images/logo.png
runtime: shiny_prerendered
output:
  learnr::tutorial:
    css: css/custom-learnr.css
subtitle: Session 2.2
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

![](https://youtu.be/nSnnTeWGW-s)

It's been said that global health is more a bunch of problems than a discipline. If you pick up a recent article on a global health topic, chances are good that you won't find much, if any, reference to theory. Instead, black box evaluations are the norm. These studies ask "does my intervention work, and to what degree", but not why, why not, or how. There are compelling reasons to think that more fully integrating theory into our work would benefit our studies in terms of design and inference, while also advancing the progress of science. In this session we'll explore how a good conceptual framework can clarify the theoretical motivation underlying your intervention. One common form of conceptual model in the evaluation literature is called a theory of change. A theory of change lifts the lid on the black box and states how intervention activities are hypothesized to generate outcomes, and what assumptions must hold for this to happen. Another way to draw these assumptions is called a logic model. We'll practice designing both models.

### Learning Objectives

* Understand what we mean by "black box evaluations"
* Explain the potential value of conceptual models
* Develop a theory of change and logic model

## Before class

### Prepare

* Read [GHR Chapter 6](https://read.themethodsection.com/theory.html){target="_blank"}.
* Patil, S. R., Arnold, B. F., Salvatore, A. L., Briceno, B., Ganguly, S., Colford Jr, J. M., & Gertler, P. J. (2014). The effect of India's total sanitation campaign on defecation behaviors and child health in rural Madhya Pradesh: a cluster randomized controlled trial. [*PLoS Medicine, 11(8), e1001709*](https://drive.google.com/open?id=0Bxn_jkXZ1lxub0xMNGJaVXNxZk0){target="_blank"}.

### Pre-Class Assessment

```{r pca02-2, echo=FALSE}
quiz(
question("In a descriptive research proposal you would not expect to see ______.",
  answer("Research problem", message = "Every study begins with a motivating research problem."),
  answer("Hypothesis testing", correct = TRUE),
 answer("Outcome", message = "Descriptive research describes specific variables in relation to an outcome"),
  allow_retry = TRUE
  ),
question("________ can take any format, so long as it articulates how an intervention, policy, or program is expected to impact an outcome. It conveys the proposed mechanism from activities to results.",
  answer("Theory of change", correct = TRUE),
  answer("Directed Acyclic Graph (DAG)", message = "DAGs are conceptual models of a system that assign causal relationships among variables and assess for potential problems."),
  answer("Models", message = "Models can serve as a bridge from a research question to measurement."),
  allow_retry = TRUE
  ),
question("To identify potential confounding variables that need to be addressed we must use _______.",
  answer("Epidemic model", message = "Epidemic models are used to explain or predict the spread of an epidemic."),
  answer("Directed acyclic graph", correct = TRUE),
  answer("Logic model", message = "Logic models are tools for program planning, monitoring program implementation, and program evaluation and reporting."),
  allow_retry = TRUE
  ),
question("______ is similar to a theory of change, but is typically more detailed and better suited for reporting progress.",
  answer("Casual impact", message = "Causal impact is the difference in counterfactual outcomes."),
  answer("Logic Model", correct = TRUE),
  answer("Logistic model", message = "Logistic model is used to model the probability of a certain event or outcome."),
  allow_retry = TRUE
  ),
question(" In Malawi, a national AIDS Control Program within the Ministry of Health was implemented. The  project team conducted awareness campaigns and distributed condoms at 19 clinical sites. When developing a logic model, the number of distributed condoms must be considered as ______ .",
  answer("Inputs", message = "Inputs in this example would be funding, staff, facilities supplies ...etc"),
  answer("Outputs", correct = TRUE),
  answer("Outcomes", message = "Outcomes in this example would be increased knowledge, decreased risky behavior ...etc"),
  answer("Impact", message = "The impact in this example would be changes in HIV prevalence, AIDS- related mortality ...etc"),
  allow_retry = TRUE
  )
)
```




<!--
<iframe height="800" width="100%" src="https://duke.qualtrics.com/jfe/form/SV_bBg1PTRLi5Usjat" frameborder="0"></iframe>
-->

## During class

### Mini Lecture

[Deck](https://docs.google.com/presentation/d/e/2PACX-1vQRfi-MYknhQWFJ_w7xwNqDkqd77pxUod0Iu6PIXAx5L-3GU_qdkhEWbu3igN27hTQ2wb94KJwMu_Fd/pub?start=false&loop=false&delayms=15000){target="_blank"}

### Activity

In this activity you will design a theory of change and logic model for the Total Sanitation Campaign.

> Poor sanitation is thought to be a major cause of enteric infections among young children. However, there are no previously published randomized trials to measure the health impacts of large-scale sanitation programs. India’s Total Sanitation Campaign (TSC) is one such program that seeks to end the practice of open defecation by changing social norms and behaviors, and providing technical support and financial subsidies. The objective of this study was to measure the effect of the TSC implemented with capacity building support from the World Bank’s Water and Sanitation Program in Madhya Pradesh on availability of individual household latrines (IHLs), defecation behaviors, and child health (diarrhea, highly credible gastrointestinal illness [HCGI], parasitic infections, anemia, growth).

<iframe src="https://drive.google.com/file/d/1b1DS0JsuNPKDnfb3viJCUBSDECZZVwtf/preview" width="100%" height="480"></iframe>

[Download activity sheets](https://drive.google.com/open?id=1VKNGv6buP9iGLPEbrbdYsw1YtcKGNkdE)

## After class

An optional activity is to review [this "how-to" note from USAID](https://usaidlearninglab.org/sites/default/files/resource/files/project_logic_model_how_to_note_final_sep1.pdf){target="_blank"} on developing a project logic model. It's smart to understand how donors like USAID think about the role and design of these conceptual models. 

If you think USAID funded research might be in your future (the answer is probably yes if you want to work at a non-governmental organization like [FHI360](https://www.fhi360.org/){target="_blank"} or a similar organization), or if you are considering applying for a job or fellowship with USAID, you should consider exploring USAID's [CLA toolkit](https://usaidlearninglab.org/cla-toolkit){target="_blank"}. CLA stands for Collaborating, Learning and Adapting. The more you know about approaches to monitoring and evaluation the better.