---
title: "Randomization"
output: 
  learnr::tutorial:
    css: css/custom-learnr.css
    progressive: true
    allow_skip: true
resource_files:
- tutorial_randomization.Rmd
- css/custom-learnr.css
- images/logo_dark.png
- images/logo.png
- images/rand.png
df_print: default
runtime: shiny_prerendered
subtitle: Tutorial
---


## Introduction

This tutorial will show you how to use the [`randomizer`](https://declaredesign.org/r/randomizr/){target="_blank"} and [`blockTools`](http://www.ryantmoore.org/html/software.blockTools.html){target="_blank"} packages for R to randomly assign units like people and villages to study arms.

## Generate fake data {data-progressive=TRUE}

To get started, we need some data. Imagine that you are working with a local NGO to prospectively evaluate one of their programs. At your request, they visit 26 villages (clusters) and enroll 1000 adults between the ages of 18 and 35 to participate in their program. Here's the enrollment data they send you.

```{r setup}
# load
  library(learnr)
  library(tidyverse)
  library(randomizr)
  library(randomNames)
  library(blockTools)
  tutorial_options(exercise.timelimit = 60)
  
# setup
  set.seed(371702) # makes this reproducible
  N <- 1000        # sample size

# create data
  enrolled <-
  data.frame(partID = seq(from=1, to=N, by=1)) %>%          # fake participants
  mutate(clusterID = sample(letters[1:26], N, replace=TRUE),# village IDs
         clusterPop = sample(200:20000, N, replace=TRUE),   # village population
         female = sample(0:1, N, replace=TRUE),             # female=1, male=0
         married = sample(0:1, N, replace=TRUE),            # married-1, not=0
         age = sample(18:35, N, replace=TRUE),              # participant age
         name = randomNames(N, gender=female)               # names
         ) %>%
  select(partID, name, female, married, age, clusterID, clusterPop)
  enrolled
```

## Randomize individuals {data-progressive=TRUE}

We we can use the `complete_ra()` function from the `randomizer` package to randomly assign individuals to study arms. In the example below we specify 2 arms labeled "treatment" and "control". Run the code and observe the counts of participants assigned to each arm.

```{r individuals, exercise=TRUE, exercise.eval=FALSE}
# randomize
  rand_ind <- 
  enrolled %>%
    select(partID, name) %>%
    mutate(arm = complete_ra(N = N,                         # N is sample size
                             num_arms = 2,
                             conditions=c("control", "treatment"))
           )
  rand_ind

# check
  rand_ind %>%
    group_by(arm) %>%
    count()
```

Try setting `num_arms` to 3, add a third `condition` label, and re-run the code.

## Block randomization {data-progressive=TRUE}

### Single blocking variable {data-progressive=TRUE}

If we want to make sure that participant sex is equally distributed across arms (more of a concern when the sample size is small), or if we want to setup a heterogeneity analysis where we look at the differential impact of the intervention by sex, we can block the randomization on sex using the `block_ra()` function. Run the code.

```{r individuals-block, exercise=TRUE, exercise.eval=FALSE}
# randomize
  rand_ind_bl <- 
  enrolled %>%
    select(partID, name, female) %>%
    mutate(arm = block_ra(blocks = female,
                          num_arms = 2,
                          conditions=c("control", "treatment"))
           )
  rand_ind_bl

# check
  rand_ind_bl %>%
    group_by(arm, female) %>%
    count()
```

### Multiple blocking variables {data-progressive=TRUE}

If we have multiple binary or categorical variables that we want to block on, we could create a variable that represents the combination of these variables and pass that to the `block_ra()` function. For instance, here we create a variable called `femaleMarried` that indicates the sex and marital status of each person. Run the code.

```{r individuals-block-multi, exercise=TRUE, exercise.eval=FALSE}
# randomize
  rand_ind_bl_m <- 
  enrolled %>%
    select(partID, name, female, married) %>%
    mutate(femaleMarried = ifelse(female==1 & married==1, "married, female",
                           ifelse(female==1 & married==0, "not married, female",
                           ifelse(female==0 & married==1, "married, male",
                           ifelse(female==0 & married==0, "not married, male",
                                  "??"))))
           ) %>%
    mutate(arm = block_ra(blocks = femaleMarried,
                          num_arms = 2,
                          conditions=c("control", "treatment"))
           )
  rand_ind_bl_m

# check
  rand_ind_bl_m %>%
    group_by(arm, femaleMarried) %>%
    count()
```

Note how our categories appear to be balanced across arms.

### Multiple blocking variables, `blockTools` {data-progressive=TRUE}

That's simple, but what if we wanted to block on multiple variables, including some continuous like `age`? One solution is to use the `blockTools` package to construct matched blocks and pass this blocking information to the `block_ra()` function. Run the code.

```{r blocktools, exercise=TRUE, exercise.eval=FALSE}
# create blocks using blockTools
  block.out <- block(data = enrolled, n.tr = 2, id.vars ="partID", 
                     algorithm="randGreedy",
                     block.vars = c("female", "age", "married"), 
                     verbose=FALSE) 
  
# conduct randomization
  rand_ind_bl_bt <-
  enrolled %>%
    select(partID, name, female, age, married) %>%
    mutate(block = createBlockIDs(block.out, ., id.var = "partID")) %>%
    mutate(arm = block_ra(blocks = block,
                          num_arms = 2,
                          conditions=c("control", "treatment"))
           )
  rand_ind_bl_bt

# check
  rand_ind_bl_bt %>%
    group_by(arm) %>%
    summarize(ageMean = mean(age), 
              femaleP = mean(female),
              marriedP = mean(married))
```

We appear to get pretty balanced arms on sex, age, and marital status.

## Randomize clusters {data-progressive=TRUE}

### No blocking {data-progressive=TRUE}

Let's say we wanted to randomize villages (clusters), not individuals. We can use the `cluster_ra()` function instead and pass the `clusterID`. Run the code.

```{r clusters, exercise=TRUE, exercise.eval=FALSE}
# randomize
  rand_cluster <- 
  enrolled %>%
    select(partID, name, clusterID) %>%
    mutate(arm = cluster_ra(clusters = clusterID,       # cluster is cluster ID
                             num_arms = 2,
                             conditions=c("control", "treatment"))
           )
  rand_cluster

# check
  rand_cluster %>%
    distinct(clusterID, .keep_all = TRUE) %>%
    group_by(arm) %>%
    count()
```

The default is to assign clusters 1:1 to the arms.

### With blocking {data-progressive=TRUE}

We can randomize by clusters and include blocking information using the `block_and_cluster_ra` function. Try running the code.

```{r clusters-bl, exercise=TRUE, exercise.eval=FALSE}
# identify cluster blocking
  cluster_block <- data.frame(clusterID = letters[1:26],
                               clusterHigh = sample(0:1, 26, replace=TRUE,
                                                    prob = c(10/26, 16/26)))
# randomize
  rand_cluster_bl <- 
  cluster_block %>%
    mutate(clusterHigh = sample(0:1, 26, replace=TRUE),
           arm = block_and_cluster_ra(clusters = clusterID, # cluster is cluster ID
                                      blocks = clusterHigh, # clusterHigh is block
                            num_arms = 2,
                            conditions=c("control", "treatment"))
           )
  rand_cluster_bl
  
# check
  rand_cluster_bl %>%
    group_by(arm, clusterHigh) %>%
    count()
```

Here we see `clusterHigh` is balanced across arms.

## To learn more {data-progressive=TRUE}

Check out the full suite of tools at [Declare Design](https://declaredesign.org){target="_blank"}. If you need to prepare a random assignment materials in advance for on-the-fly assignment, have a look at the [Randomizer](https://www.randomizer.org/tutorial/){target="_blank"} web app.